Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-06-28 Documents/Paper/ms.tex
Adding extra details

\section{Introduction}
\setlength{\parindent}{0cm}
\setlength{\parskip}{3mm}
Many inference problems have the following structure. There is some unknown
number, $N$, of objects in a region. Each of the $N$ objects has a property
$\xx$, which may be a single scalar value (for example, a mass), or a set of
values (e.g. a position and a mass).

@@

The model for the signal is
\begin{eqnarray}
y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
\end{eqnarray}
where there are $N$ sinusoids in the signal, the
amplitudes are $\{A_i\}$, the periods are $\{T_i\}$, and the phases are
$\{\phi_i\}$.

The sampling distribution (probability distribution for the data given the
parameters) is a normal (gaussian) distribution with mean zero and standard
deviation $\sigma$, applied independently to each data point:
\begin{eqnarray}
Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\} \sim
\mathcal{N}\left(\mu(t_i), \sigma^2\right).
\end{eqnarray}
We simulated some data based on the assumption $N=2$.
The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a
large, low period oscillation with a smaller, much faster oscillation
superimposed. The noise level is such that the fast oscillation is difficult
to detect. The true values of the parameters were
$N=2$, $\mathbf{A} = \{1, 0.3\}$,
$\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
observed at $n=1001$ points equally spaced between $t=0$ and $t=100$.
The true form of the signal is:
\begin{eqnarray}
y(t) = \sin\left(\frac{2\pi t}{30}\right) +
0.3 \sin\left(\frac{2\pi t}{2} + 1\right)
\end{eqnarray}
and the noise standard deviation was $\sigma = 1$. The data $\{Y_i\}_{i=1}^n$
is displayed in Figure~\ref{fig:sinewave_data}.

>>>>

\section{Introduction}\label{sec:introduction}
\setlength{\parindent}{0cm}
\setlength{\parskip}{3mm}
Many inference problems have the following structure. There is some unknown
number, $N$, of objects in a region. Each of the $N$ objects has a property
$\xx$, which may be a single scalar value (for example, a mass), or a set of
values (e.g. a position and a mass).

@@

In the notation of Section~\ref{sec:introduction}, the parameters of the
each object are
\begin{eqnarray}
\xx_i &=& \{A_i, T_i, \phi_i\}.
\end{eqnarray}

The model for the shape of the (noise-free) signal is
\begin{eqnarray}
y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
\end{eqnarray}
where there are $N$ sinusoids in the signal, the
amplitudes are $\{A_i\}$, the periods are $\{T_i\}$, and the phases are
$\{\phi_i\}$.
The sampling distribution (probability distribution for the data given the
parameters) is a normal (gaussian) distribution with mean zero and standard
deviation $\sigma$, applied independently to each data point:
\begin{eqnarray}
Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\}, \sigma \sim
\mathcal{N}\left(\mu(t_i), \sigma^2\right).
\end{eqnarray}
This also depends on an additional noise standard deviation parameter $\sigma$
which will also be inferred.

To demonstrate the techniques,
we simulated some data based on the assumption $N=2$, i.e. the true signal was
composed of two sinusoids.
The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a
large, low period oscillation with a smaller, much faster oscillation
superimposed. The noise level is such that the fast oscillation is difficult
to detect. The true values of the parameters were
$N=2$, $\mathbf{A} = \{1, 0.3\}$,
$\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
observed at $n=1001$ points equally spaced between $t=0$ and $t=100$.
The true form of the signal is:
\begin{eqnarray}
y(t) = \sin\left(\frac{2\pi t}{30}\right) +
0.3 \sin\left(\frac{2\pi t}{2} + 1\right)
\end{eqnarray}
and the noise standard deviation was $\sigma = 1$. The data $\{Y_i\}_{i=1}^n$
is displayed in Figure~\ref{fig:sinewave_data}.
